---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# toxdrc

<!-- badges: start -->
<!-- badges: end -->

The goal of toxdrc is to provide a simple, easy solution to model and estimate the dose response relationships in complex data sets. This package provides a suite of modular functions that can take raw data, perform quality checks, apply user-specified transformation/normalization steps, fit and select the best fitting dose-response model, and estimate specified effect measures across multiple experimental subsets.

## Installation

You can install the development version of toxdrc from [GitHub](https://github.com/jsalole/toxdrc) with:

``` r
# install.packages("pak")
pak::pak("jsalole/toxdrc")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(toxdrc)
print(cellglow)
```

At first glance this dataset seems manageable, but it consists of 45 subsets of data. 

Current approaches would require us to fragment the dataset and analyze each one manually.

`toxdrc` offers another way to do this, through `runtoxdrc()`, which analyzes all of them in one function.

```{r}
analyzed_data <- runtoxdrc(
  dataset = cellglow,
  Conc = Conc,
  Response = RFU,
  IDcols = c("Test_Number", "Dye", "Replicate", "Type"),
  quiet = TRUE,
  qc = toxdrc_qc(), # uses defualt arguments, which can be overwritten. This controls testing for solvent effects, outlier testing, and CV calculations. Use ?toxdrc_qc() for more info.
  normalization = toxdrc_normalization(
    # here we specify to blank correct and normalize, instead of using default values. Only have to specify the arguments that do not use the default value. Use ?toxdrc_normalization() for more info.
    blank.correction = TRUE,
    normalize.resp = TRUE
  ),
  toxicity = toxdrc_toxicity(), # uses defualt arguments, which can be overwritten. This controls how toxicity is determined. Use ?toxdrc_toxicity() for more info.
  modelling = toxdrc_modelling(
    # Use ?toxdrc_modelling() for more info.
    quiet = TRUE
  )
)

analyzed_data
```

Currently working on other functions to view the plots and combine the output into useable, readable formats.