---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# toxdrc

<!-- badges: start -->
[![R-CMD-check](https://github.com/jsalole/toxdrc/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jsalole/toxdrc/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**toxdrc** provides a streamlined workflow for analyzing dose-response data.  
The package includes a set of modular functions that collectively:

- import and clean raw experimental data
- run quality checks
- apply optional user-defined transformations or normalizations
- fit multiple dose–response models
- automatically select the best-fitting model
- estimate specified effect measures across experimental groups or conditions

This makes **toxdrc** well-suited for handling complex or multi-factor experiments where dose–response relationships must be modelled consistently.

## Getting Started

You can install the development version of **toxdrc** from [GitHub](https://github.com/jsalole/toxdrc) with:

``` r
# install.packages("pak")
pak::pak("jsalole/toxdrc")
```

```{r}
library("toxdrc")
```

This will get the package into your R session. 

## Examples

This package comes built in with two example datasets. `toxresult` is a small dataset that looks at the results of a toxicity test within a single 24-well plate. `cellglow` is much larger with 1080 observations. These demonstrate when this package is useful.

```{r}
length(unique(interaction(toxresult$TestID, toxresult$Dye)))
```

```{r}
length(unique(interaction(cellglow$TestID, cellglow$Dye)))
```

With one  set of data, existing options like the `drc` package can be quite manageable. These can also be assessed using **toxdrc** through the many modular functions that exist, like `blankcorrect()`, `normalizeresponse()`, `modelcomp()`, `getECx()`, among many others. This package is most useful is through the pipeline function, `runtoxdrc()`. This pipeline splits up the dataset into the subsets for analysis.

```{r}
data_01 <- runtoxdrc(
  dataset = cellglow,
  Conc = Conc,
  Response = RFU,
  IDcols = c("Test_Number", "Dye", "Replicate", "Type"),
  quiet = TRUE,
  qc = toxdrc_qc(),
  normalization = toxdrc_normalization(
    blank.correction = TRUE,
    normalize.resp = TRUE
  ),
  toxicity = toxdrc_toxicity(),
  modelling = toxdrc_modelling(
    quiet = TRUE,
  ),
  output = toxdrc_output()
)
```

```{r}
data_02 <- runtoxdrc(
  dataset = cellglow,
  Conc = Conc,
  Response = RFU,
  IDcols = c("Test_Number", "Dye", "Replicate", "Type"),
  quiet = TRUE,
  qc = toxdrc_qc(),
  normalization = toxdrc_normalization(
    blank.correction = TRUE,
    normalize.resp = TRUE
  ),
  toxicity = toxdrc_toxicity(),
  modelling = toxdrc_modelling(
    quiet = TRUE,
  ),
  output = toxdrc_output(
    condense = TRUE
  )
)
```

While the pipeline only requires a few functions, it can be tailored to meet the needs of most assays. The `toxdrc_config()` help file indicates how these arguments can be used.

The `data_01` provides intermediate information on the pipeline for each subset, but `data_02` will return a simple dataframe summarizung the results.

```{r}
length(data_01)
str(data_01[1])
```

```{r}
data_02
```

This package is also capable of making multiple point-estimates for each subset:

```{r}
data_03 <- runtoxdrc(
  dataset = cellglow,
  Conc = Conc,
  Response = RFU,
  IDcols = c("Test_Number", "Dye", "Replicate", "Type"),
  quiet = TRUE,
  normalization = toxdrc_normalization(
    blank.correction = TRUE,
    normalize.resp = TRUE
  ),
  modelling = toxdrc_modelling(
    EDx = c(0.2, 0.5, 0.7),
    quiet = TRUE
  ),
  output = toxdrc_output(condense = TRUE)
)

head(data_03)
```

## Authours
* **Jack Salole** - *Intial Work* - [jsalole](https://github.com/jsalole) 

## Acknowledgments
* This package is an extenision of the [drc package](https://cran.r-project.org/web/packages/drc/drc.pdf).
* This package uses code suggested by [Nel on stack overflow](https://stackoverflow.com/users/7133643/nel) for mselect2, which updates mselect to work within a function.
